Do we need detailed demographic data to forecast plant population responses to climate change?

### 2016-07-06 manuscript/manuscript.Rmd
updates ms with minor edits from MBH

### Stastical models of vital rates
At both levels of inference (individual and quadrat), the building blocks of our population models are vital rate regressions. For individual-level data, we fit regressions for survival, growth, and recruitment for each species. At the quadrat-level, we fit a single regression model for population growth. We describe the statistical models separately because they required different approaches. For both model types, we fit vital rate models with and without climate covariates. Models with climate effects contain five climate covariates that we chose \emph{a priori} based on previous model selection efforts using these data [@Chu2016] and expert advice (Lance Vermeire, *personal communication*): "water year" precipitation at \emph{t}-2 (lagppt); April through June precipitation at \emph{t}-1 and \emph{t} (ppt1 and ppt2, respectively) and April through June temperature at \emph{t}-1 and \emph{t} (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t}-1 to *t* is the transition of interest. We also include interactions among same-year climate covariates (e.g., ppt1 $\times$ TmeansSpr1), resulting in a total of seven climate covariates.

@@

\vspace{-3em}
\begin{align}
y_{i,q,t+1}^G &\sim \text{Normal}(\mu_{i,q,t+1}, \sigma^2_{x_{i,q,t+1}}\textbf{I}), \\
\mu_{i,q,t+1} &= \beta_{0,t} + \beta_{s,t}x_{i,q,t} + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c + \beta_{d,1} w_{i,q,t} + \beta_{d,2} (x_{i,q,t}w_{i,q,t}),
\end{align}
\vspace{-3em}

@@

To find the optimal penalty (i.e., optimal value of the hyperparameter $\sigma_{\beta_c}^2$), we fit each statistical model with a range of values for $\sigma_{\beta_c}^2$ and compared predictive scores from leave-one-year-out cross-validation.
We performed the grid search over 24 values of $\sigma_{\beta_c}^2$, ranging from $\sigma_{\beta_c}^2 = 0.01$ to $\sigma_{\beta_c}^2 = 2.25$.
For each statistical model and each species, we fit $13\times24=312$ iterations of the model fitting algorithm to search $\sigma_{\beta_c}^2$ for the optimal value (13 years to leave out for cross-validation and 24 values of $\sigma_{\beta_c}^2$) -- a total of 4,992 model fits.
We calculated the log pointwise predictive density (*lppd*) to score each model's ability to predict the left-out data [@Gelman2014].
Thus, for training data $y_{train}$ and held-out data $y_{hold}$ at a given value of  $\sigma_\theta^2$ across all MCMC samples $s=1,2,...,S$ and all hold outs of data from year *t* to year *T*, and letting $\theta$ represent all unknowns, *lppd* is

\vspace{-3em}
\begin{align}
\text{lppd}_{\text{CV}} = \sum_{t=1}^T \text{log}_e \int [y_{t,hold}|\theta][\theta|y_{train}]d\theta,
\end{align}
\vspace{-3em}

@@

\vspace{-3em}
\begin{align}
\sum_{t=1}^T \text{log}_e \left(\frac{1}{S} \sum^S_{s=1}(y_{t,hold}|\theta_{ts})  \right).
\end{align}
\vspace{-3em}

@@

We used random draws from the final 1,000 iterations from each of three MCMC chains for each vital rate regression to carry-through parameter uncertainty into our population models. At each time step, we drew the full parameter set (climate effects and density-dependence fixed effects) from a randomly selected MCMC iteration. Relatively unimportant climate covariates (those that broadly overlap 0) will have little effect on the mean of the simulation results, but can contribute to their variation. To retain temporal variation associated with random year effects, we used posterior estimates of the mean temporal effect and the standard deviation of that effect to generate a random year effect for unobserved years. That is, for some future year *T*, the intercept is $\beta_0,T \sim \text{Normal}(\beta_0, \sigma_{\beta_0})$ and the effect of size is $\beta_s,T \sim \text{Normal}(\beta_s, \sigma_{\beta_s})$.

>>>>

### Statistical models of vital rates
At both levels of inference (individual and quadrat), the building blocks of our population models are vital rate regressions. For individual-level data, we fit regressions for survival, growth, and recruitment for each species. At the quadrat-level, we fit a single regression model for population growth. We describe the statistical models separately because they required different approaches. For both model types, we fit vital rate models with and without climate covariates. Models with climate effects contain five climate covariates that we chose \emph{a priori} based on previous model selection efforts using these data [@Chu2016] and expert advice (Lance Vermeire, *personal communication*): "water year" precipitation at \emph{t}-2 (lagppt); April through June precipitation at \emph{t}-1 and \emph{t} (ppt1 and ppt2, respectively) and April through June temperature at \emph{t}-1 and \emph{t} (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t}-1 to *t* is the transition of interest. We also include interactions among same-year climate covariates (e.g., ppt1 $\times$ TmeansSpr1), resulting in a total of seven climate covariates.

@@

\vspace{-3em}
\begin{align}
y_{i,q,t+1}^G &\sim \text{Normal}(\mu_{i,q,t+1}, \sigma^2_{x_{i,q,t+1}}), \\
\mu_{i,q,t+1} &= \beta_{0,t} + \beta_{s,t}x_{i,q,t} + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c + \beta_{d,1} w_{i,q,t} + \beta_{d,2} (x_{i,q,t}w_{i,q,t}),
\end{align}
\vspace{-3em}

@@

To find the optimal penalty (i.e., optimal value of the hyperparameter $\sigma_{\beta_c}^2$), we fit each statistical model with a range of values for $\sigma_{\beta_c}^2$ and compared predictive scores from leave-one-year-out cross-validation.
We performed the grid search over 24 values of $\sigma_{\beta_c}^2$, ranging from $\sigma_{\beta_c}^2 = 0.01$ to $\sigma_{\beta_c}^2 = 2.25$.
For each statistical model and each species, we fit $13\times24=312$ iterations of the model fitting algorithm to search $\sigma_{\beta_c}^2$ for the optimal value (13 years to leave out for cross-validation and 24 values of $\sigma_{\beta_c}^2$) -- a total of 4,992 model fits.
We calculated the log pointwise predictive density (*lppd*) to score each model's ability to predict the left-out data [@Gelman2014].
Thus, for training data $y_{\text{train}}$ and held-out data $y_{\text{hold}}$ at a given value of  $\sigma_\theta^2$ across all MCMC samples $s=1,2,...,S$ and all hold outs of data from year *t* to year *T*, and letting $\theta$ represent all unknowns, *lppd* is

\vspace{-3em}
\begin{align}
\text{lppd}_{\text{CV}} = \sum_{t=1}^T \text{log}_e \int [y_{t,\text{hold}}|\theta][\theta|y_{\text{train}}]d\theta,
\end{align}
\vspace{-3em}

@@

\vspace{-3em}
\begin{align}
\sum_{t=1}^T \text{log}_e \left(\frac{1}{S} \sum^S_{s=1}(y_{t,\text{hold}}|\theta_{ts})  \right).
\end{align}
\vspace{-3em}

@@

We used random draws from the final 1,000 iterations from each of three MCMC chains for each vital rate regression to carry-through parameter uncertainty into our population models. At each time step, we drew the full parameter set (climate effects and density-dependence fixed effects) from a randomly selected MCMC iteration. Relatively unimportant climate covariates (those that broadly overlap 0) will have little effect on the mean of the simulation results, but can contribute to their variation. To retain temporal variation associated with random year effects, we used posterior estimates of the mean temporal effect and the standard deviation of that effect to generate a random year effect for unobserved years. That is, for some future year *T*, the intercept is $\beta_{0,T} \sim \text{Normal}(\beta_0, \sigma_{\beta_0})$ and the effect of size is $\beta_{s,T} \sim \text{Normal}(\beta_s, \sigma_{\beta_s})$.
